---
title: 'Group Project: COVID-19 Responses'
author: "Julie Trundle"
date: "2/15/2022"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(here)
library(dplyr)
library(knitr)
library(naniar)
library(shiny)
library(shinydashboard)
library(lubridate)
library(ggVennDiagram)
```

```{r}
covid <- read_csv(here("Gov_Responses2Covid19_last (1).csv"))
```

```{r}
view(covid)
```

```{r}
summary(covid)
```

```{r}
glimpse(covid)
```

```{r}
naniar::miss_var_summary(covid)
```
```{r}
dim(covid)
```

```{r}
names(covid)
```

```{r}
covid <- covid %>% 
  rename(cov_date = d)
names(covid)
```

```{r}
covid_clean <- covid %>% 
  mutate(cov_date = dmy(cov_date)) %>% 
  clean_names()
covid_clean
```

Total number of cases per country: 
```{r}
covid_clean %>% 
  select(country, cases) %>% 
  group_by(country) %>% 
  summarise(sum_cases=sum(cases)) %>% 
  arrange(desc(sum_cases))
```

```{r}
covid_clean %>% 
  group_by(country) %>% 
  summarise(sum_cases=sum(cases)) %>% 
  top_n(10, sum_cases) %>% 
  ggplot(aes(x=country, y=sum_cases, fill=country))+
  geom_col()+
  theme(axis.text.x = element_text(angle=30))
```

Top 5 countries with the most total covid cases over 2020: United States, India, Brazil, Russia, Colombia. 

Total sum of deaths per country: 
```{r}
covid_clean %>%
  select(country, deaths) %>%
  group_by(country) %>%
  summarise(sum_deaths=sum(deaths)) %>%
  arrange(desc(sum_deaths))
```
```{r}
covid_clean %>% 
  group_by(country) %>% 
  summarise(sum_deaths=sum(deaths)) %>% 
  top_n(10, sum_deaths) %>% 
  ggplot(aes(x=country, y=sum_deaths, fill=country))+
  geom_col()+
  theme(axis.text.x = element_text(angle=30))
```
Top 5 countries with the most total deaths over 2020: United States of America, Brazil, India, Mexico, United Kingdom. 

Facet wrap with cases of the top 5 countries with most cases: 

```{r}
covid_clean %>% 
  filter(country=="United States of America" | country=="India" | country=="Brazil" | country=="Russia" | country=="Colombia") %>% 
  select(country, cov_date, cases) %>% 
  group_by(country) %>% 
  summarize(sum_cases = sum(cases))
```


```{r}
covid_clean %>% 
  select(country, cov_date, cases) %>% 
  filter(country=="United States of America" | country=="India" | country=="Brazil" | country=="Russia" | country=="Colombia") %>%
  ggplot(aes(x=cov_date, y=cases, color=country))+
  scale_y_log10()+
  geom_point(size=0.2)+
  geom_line()+
  facet_wrap(~country)+
  theme(axis.text.x = element_text(size=0))+
  labs(title = "Top 5 Countries in Number of Cases",
       x = "Date",
       y = "Cases")
```


Economic measures taken by each country: 
```{r}
covid_clean %>% 
  select(country, economic_measures) %>% 
  filter(economic_measures!="NA") %>% 
  group_by(country) %>% 
  summarize(mean_economic_measures=mean(economic_measures)) %>% 
  arrange(desc(mean_economic_measures))
```
```{r}
covid_clean %>% 
  select(country, economic_measures) %>% 
  filter(economic_measures!="NA") %>% 
  group_by(country) %>% 
  summarize(mean_economic_measures=mean(economic_measures)) %>% 
  arrange(mean_economic_measures)
```

Looking at difference in percent of cases between Thailand (most economic measures) and Eritrea (an example of no economic measures)
```{r}
covid_clean %>% 
  select(country, cov_date, economic_measures, cases, population_2019) %>% 
  filter(country=="Thailand" | country=="Eritrea") %>% 
  mutate(percent_cases = cases/population_2019) %>% 
  ggplot(aes(x=cov_date, y=percent_cases, color=country))+
  geom_point(size=0.8)+
  geom_line()+
  geom_smooth(method=lm, se=F)+
  theme(axis.text.x = element_blank())+
  labs(title = "Percent Cases Comparison: Thailand and Eritrea",
       x = "Date",
       y = "Percent of Cases")
```
```{r}
covid_clean %>% 
  select(country, rigidity_public_health) %>% 
  filter(rigidity_public_health!="NA") %>% 
  group_by(country) %>%
  summarize(mean_public_health_measures=mean(rigidity_public_health)) %>% 
  arrange(desc(mean_public_health_measures))
```
```{r}
covid_clean %>% 
  select(country, rigidity_public_health) %>% 
  filter(rigidity_public_health!="NA") %>% 
  group_by(country) %>%
  summarize(mean_public_health_measures=mean(rigidity_public_health)) %>% 
  arrange((mean_public_health_measures))
```

```{r}
covid_clean %>% 
  select(country, cov_date, rigidity_public_health, cases, population_2019) %>% 
  filter(country=="Morocco" | country=="Monaco") %>% 
  mutate(percent_cases = cases/population_2019) %>% 
  ggplot(aes(x=cov_date, y=percent_cases, color=country))+
  geom_point(size=0.8)+
  geom_line()+
  geom_smooth(method=lm, se=F)+
  theme(axis.text.x = element_blank())+
  labs(title = "Percent Cases Comparison: Morocco and Monaco",
       x = "Date",
       y = "Percent of Cases")
```

```{r}
covid_clean %>% 
  select(country, cov_date, economic_measures) %>% 
  filter(country=="United States of America" | country=="India" | country=="Brazil" | country=="Russia" | country=="Colombia") %>%
  ggplot(aes(x=cov_date, y=economic_measures, color=country))+
  geom_smooth(size=0.6, se=F)+
  theme(axis.text.x = element_text(size=0))+
  labs(title = "Economic Measures of ")
```
Scatter plot showing economic measures vs public health measures:

```{r}
covid_clean %>% 
  select(country, economic_measures, rigidity_public_health) %>% 
  filter(economic_measures!="NA") %>% 
  filter(rigidity_public_health!="NA") %>% 
  group_by(country) %>% 
  summarize(mean_economic_measures=mean(economic_measures),
            mean_public_health_measures=mean(rigidity_public_health)) %>% 
  arrange(desc(mean_economic_measures))
```

```{r}
covid_clean %>% 
  select(country, economic_measures, rigidity_public_health) %>% 
  filter(economic_measures!="NA") %>% 
  filter(rigidity_public_health!="NA") %>% 
  group_by(country) %>% 
  summarize(mean_economic_measures=mean(economic_measures),
            mean_public_health_measures=mean(rigidity_public_health)) %>% 
  ggplot(aes(x=mean_economic_measures, y=mean_public_health_measures))+
  geom_point(size=1)+
  geom_smooth(method=lm)+
  labs(title = "Economic Measures vs. Public Health Measures",
       x = "Mean Economic Measures",
       y = "Mean Public Health Measures")+
  theme(title = element_text(size=14),
        axis.title = element_text(size=11),
        axis.text = element_text(size=8))
```

Venn Diagram showing countries with more than average economic measures vs public health measures: 

```{r}
summary(covid_clean)
```

```{r}
economic_measures_implemented <- covid_clean %>% 
  filter(economic_measures>="0.385") %>% 
  pull(country)

public_health_measures_implemented <- covid_clean %>% 
  filter(rigidity_public_health>="0.301") %>% 
  pull(country)
```

```{r}
high_health_measures <- list(economic_measures_implemented, public_health_measures_implemented)
```

```{r}
ggVennDiagram(high_health_measures, category.names = c("Economic Measures", "Health Measures"))
```

```{r}
economic_measures_not_implemented <- covid %>% 
  filter(Economic_Measures<="0.385") %>% 
  pull(country)

public_health_measures_not_implemented <- covid %>% 
  filter(Rigidity_Public_Health<="0.301") %>% 
  pull(country)
```

```{r}
low_health_measures <- list(economic_measures_not_implemented, public_health_measures_not_implemented)
```

```{r}
ggVennDiagram(low_health_measures, category.names = c("Low Economic Measures", "Low Public Health Measures"))
```